---
title: "04 - TakeHome"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.”
\
Kasey Wood
\
\
Use the leaflet package and the us.cities data set from the maps package to make a Zoom-able map of the 50 U.S. capital cities. When you click on a marker for a capital city, both the city name and state as well as the pop (population) should be displayed.
\
Note that, in the us.cities data set, state capitals are given a value of 2 in the capital variable.
\
A second note is that if you cannot figure out how to get two “things” (in this case name and population) to display in Leaflet, you would just lose a single point for the quiz (so don’t spend too much time on that part unless you are inclined to do so).
\
```{r}
library(tidyverse)
library(leaflet)
library(maps)
data("us.cities")
uscities_map <- us.cities %>% filter(capital == "2")
content <- uscities_map %>% mutate(popup = paste0('<a href =', uscities_map$us.cities, '>',
                        uscities_map$name,', ', 
                        uscities_map$pop))
us.cities_map <- leaflet(us.cities) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Wikimedia) %>% 
  addMarkers(lng = uscities_map$long, lat = uscities_map$lat,
             clusterOptions = markerClusterOptions(),
             popup = content$popup)
us.cities_map
```
