---
title: "05-takehome"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.”
\
Kasey Wood
\
\
Use the STAT 113 survey data set to explore the trend in the popularity of Facebook over time. Do this in 2 ways:
\
(7 points). Construct a graph that uses the number of Facebook friends each student has.

```{r}
library(tidyverse)
statsurvey_df <- read_csv("data/stat113_survey.csv")

statsurvey_fb <- statsurvey_df %>%
  filter(na.rm = TRUE) %>% 
  group_by(time_both) %>%
  summarise(nfriends = sum(Facebook, na.rm = TRUE))

statsurvey_fb <- statsurvey_fb %>% 
  separate(time_both, into = c("semester", "year"), sep = 1)

statsurvey_fb <- statsurvey_fb %>% 
  arrange(desc(year), desc(semester)) %>% 
  unite(col = "time_both", c(semester, year)) %>%
  mutate(time_both = fct_inorder(time_both))
  
ggplot(data = statsurvey_fb, aes(x = time_both, y = nfriends)) +
  geom_col() +
  coord_flip()
```

(7 points). Construct a graph that uses the proportion of students that have Facebook (assuming that 0 friends means that the student does not have a Facebook account and more than 0 friends means that the student does have a Facebook account).
```{r}
statsurvey_fb2 <- statsurvey_df %>%
  filter(na.rm = TRUE) %>%
  mutate(facebook = if_else(Facebook >= 1, "yes", "no"))

statsurvey_graph <- statsurvey_fb2 %>% 
  filter(na.rm = TRUE) %>% 
  group_by(time_both, facebook) %>%
  summarise(ncount = n()) %>%
  ungroup() %>%
  group_by(time_both) %>%
  mutate(ntotal = sum(ncount)) %>%
  ungroup() %>%
  filter(facebook == "yes") %>%
  mutate(prop = ncount / ntotal)

statsurvey_graph <- statsurvey_graph %>% 
  separate(time_both, 
           into = c("semester", "year"), 
           sep = 1)
statsurvey_graph <- statsurvey_graph %>% 
  arrange(year, desc(semester)) %>% 
  unite(col = "time_both", c(semester, year)) %>%
  mutate(time_both = fct_inorder(time_both))


ggplot(data = statsurvey_graph, aes(x = time_both, y = prop)) +
  geom_col() +
  labs(y = "proportion",
       x = "semester")
```

1 point is provided for committing and pushing to GitHub correctly.
