---
title: "6.6 Supplementary Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
hpi_df <- read_csv("data/hpi-tidy.csv")
```

Exercise 1. Make a map of a variable of your choosing. In coord_map(), use projection = "mercator", which is also the default (we will see in a later exercise that this probably is not the best choice).

```{r}
library(maps)
world_df <- ggplot2::map_data("world")
hpi_full <- left_join(world_df, hpi_df, by = c("region" = "Country"))
ggplot(data = hpi_full, 
       mapping = aes(x = long, y = lat,
                     group = group)) +
  geom_polygon(aes(fill = LifeExpectancy), colour = "black") +
  coord_map(projection = "mercator") +
  theme_void() +
  scale_fill_distiller(direction = 1)
```

Exercise 2. You may notice that the United States does not get coloured in your map. Examine this issue further and fix the map so that the United States is coloured.

```{r}
hpi_df[hpi_df$Country == "United States of America", "Country"] <- "USA"
hpi_full <- left_join(world_df, hpi_df, by = c("region" = "Country"))
ggplot(data = hpi_full, 
       mapping = aes(x = long, y = lat,
                     group = group)) +
  geom_polygon(aes(fill = LifeExpectancy), colour = "black") +
  coord_map(projection = "mercator") +
  theme_void() +
  scale_fill_distiller(direction = 1)
```

Exercise 3. You may have noticed that there are two horizontal stripes across your map. This is an issue that drove me nuts! Check out this submitted issue on ggplot2’s GitHub page for the reason for the error as well as a fix. Use it to fix your plot.

```{r}
ggplot(data = hpi_full, 
       mapping = aes(x = long, y = lat,
                     group = group)) +
  geom_polygon(aes(fill = LifeExpectancy), colour = "black") +
  coord_map(projection = "mercator", xlim = c(-180,180)) +
  theme_void() +
  scale_fill_distiller(direction = 1)
```

Exercise 4. Read about Mercator projections in this blog post. What does this source say about the sizes of Greenland vs. Africa in a Mercator projection.
\
This source says that Greenland is way bigger than it should be in comparison to Africa. The actual sizes of the countries would mean that Greenland could fit into Africa 14 times, but this map shows that Greenland is bigger or about the same size as Africa. This is because in a Mercator projection the countries farthest from the equator are represented as much larger than they actually are in comparison to the countries closer to the equator.
\
\
Exercise 5. Examine all of the different options for map projection with ?mapproject. Then, change the projection to "globular". Change the projection again to "gilbert". How does the relative size of Greenland to Africa change in the projections?
```{r}
ggplot(data = hpi_full, 
       mapping = aes(x = long, y = lat,
                     group = group)) +
  geom_polygon(aes(fill = LifeExpectancy), colour = "black") +
  coord_map(projection = "globular", xlim = c(-180,180)) +
  theme_void() +
  scale_fill_distiller(direction = 1)
## We can see that the relative size of Greenland looks much more normal in this projection than the mercator projection.
ggplot(data = hpi_full, 
       mapping = aes(x = long, y = lat,
                     group = group)) +
  geom_polygon(aes(fill = LifeExpectancy), colour = "black") +
  coord_map(projection = "gilbert", xlim = c(-180,180)) +
  theme_void() +
  scale_fill_distiller(direction = 1)
## Here we can see that the relative size of Greenland to Africa looks more normal here than the mercator projection, but still looks like a large mutation in comparison to the globular projection
```